{\bf Systematic Dynamic Race Condition Detection}

### 2012-11-01 landslide-lecture/lecture.tex
add landslide lecture and defence

\title[Landslide]{{\bf Landslide: Systematic Dynamic Race Detection in Kernel-space}}
\subtitle[]{ {\em showing that Pebbles are less stable than you thought since 2011.}}
\author[Ben Blum]{Ben Blum \texttt{(bblum@andrew.cmu.edu)}}

\institute[CMU]{Carnegie Mellon University}
\date[]{2012, May 10}

@@

\usecolortheme[named=RoyalBlue]{structure}

@@

\newcommand\linegap{\vspace{0.2in}}
\newcommand\breakslide[1]{\begin{frame}{} \begin{center} #1 \end{center} \end{frame}}
\newcommand\related[1]{\textsuperscript{\em [#1]}}

\begin{frame}{Outline}
	\textbf{Motivation: Concurrency debugging}
	\begin{itemize}
		\item Systematic testing versus stress testing
		\item Challenges of kernel-space
	\end{itemize}
	\linegap

	{\bf Tool: Landslide}
	\begin{itemize}
		\item Design and interface
		\item Addressing challenges
	\end{itemize}
	\linegap

	{\bf Evaluation: Finding Races}
	\begin{itemize}
		\item Student user study
		\item Case study
	\end{itemize}
	\linegap

	{\bf Future Work and Conclusion}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Race Conditions and the Finding Thereof}

	\begin{itemize}
		\item On exit, child's state is freed
		\item Forking thread does use-after-free
		\item Might return garbage instead of thread ID
	\end{itemize}
\end{frame}

\begin{frame}{Testing Techniques} % slide: sys ex vs stress testing
	\textbf{Stress Testing:} Common testing approach
	\begin{itemize}
		\item Attempting to exercise as many interleavings as possible
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at {\em just the right time}\dots''
		\end{itemize}
		\item Cryptic panic messages or machine reboots
	\end{itemize}
	\linegap

	{\bf Systematic Testing} \related{Musuvathi '08}
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Run {\em every single} interleaving
		\item Provide better debugging information, reproducibility
	\end{itemize}
\end{frame}

\begin{frame}{Execution Tree} % slide: decision tree
	\begin{center}
		\includegraphics[width=\textwidth]{tree.png}
	\end{center}
\end{frame}

\begin{frame}{Finding Races in Kernel-space} % slide: 15-410 6-week project why we need it in kernels
	\textbf{15-410:} Undergraduate Operating Systems
	\begin{itemize}
		\item Project 3: students write a kernel in 6 weeks.
		\item ``Pebbles'' is a UNIX-like, preemptible kernel specification.
	\end{itemize}
	\linegap

	Previous work in kernel-space has not used systematic testing.
	%\begin{itemize}
	%	\item Data race detection\related{Erickson '10}
	%	\item Failure resiliency\related{Kadav '09}
	%\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Challenges of Kernel-space}

\breakslide{Why is this environment different from all other environments?}

\begin{frame}{Causes of Concurrency} % slide: causes of concurrency, thread scheduling
	\textbf{Kernels contain their own concurrency implementation.}

	\linegap
	Context switching
	\begin{itemize}
		\item Non-deterministic timer-driven thread scheduling
	\end{itemize}
	Runqueue tracking
	\begin{itemize}
		\item Which threads are runnable?
	\end{itemize}
	Thread lifecycle tracking
	\begin{itemize}
		\item When are threads created/destroyed?
	\end{itemize}
\end{frame}

% slide: ad-hoc communication, memory conflicts (freeing, runqueue)
%\begin{frame}{Ad-Hoc Thread Communication}
%	No message-passing % TODO this slide sucks
%	
%	User-space systematic exploration relies on it\related{Simsa '10}; we cannot.
%	
%	Kernel threads share state via globals, heap.
%\end{frame}

\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	% say: it's also important to understand the kernel's internals instead of just viewing it as one ``black box''
	\begin{center}
	\includegraphics[width=\textwidth]{pebbles.png}
	\end{center}
	% now that we see these internal interfaces, how do we apply systematic testing to it?
\end{frame}
\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	``Everything interleaves with everything else''?
	\begin{itemize}
		% Say: "we don't want to have to explore every possible interaction between each of these components"
		\item Impractical to explore resulting state spaces
	\end{itemize}
	\linegap

	What is a ``bug'', anyway?\related{Engler '01}
	\begin{itemize}
		\item False-negative-oriented approach
	\end{itemize}
\end{frame}

\begin{frame}{Contribution} % slide: summary: "sys ex is a tool \ldots``
	Systematic testing is a {\em tool} for testing concurrent systems.
	\begin{itemize}
		\item How can it be applied in kernel-space?
		\item What simple things can we ask of the user to make it possible?
	\end{itemize}
	\linegap

	{\bf Landslide} is an effort to answer these questions.
	% say: landslide is an exploratory work into finding out what lies behind these questions.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Landslide}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand\done[1]{\textcolor{gray}{\em \small #1}}

%\begin{frame}{Outline}
%	\done{Motivation: Concurrency debugging}
%	\begin{itemize}
%		\item \done{Systematic exploration versus stress testing}
%		\item \done{Challenges of kernel-space}
%	\end{itemize}
%	\linegap
%
%	{\bf Tool: Landslide}
%	\begin{itemize}
%		\item {\bf Design and interface}
%		\item {\bf Addressing challenges}
%	\end{itemize}
%	\linegap
%
%	Evaluation: Finding Races
%	\begin{itemize}
%		\item Student user study
%		\item Case study
%	\end{itemize}
%	\linegap
%
%	Future Work and Conclusion
%\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Design}

\begin{frame}{Environment}
	\textbf{Simics:} a full-system x86 simulator
	\begin{itemize}
		\item Students in 15-410 use Simics to test their kernels.
		\item Landslide is a transparently-embedded Simics module.
	\end{itemize}
\end{frame}

\begin{frame}{Anatomy of Landslide}
	\begin{center}
	\includegraphics[width=0.7\textwidth]{landslide.png}
	\end{center}
	% Remember to talk about solving the thread scheduling challenge
\end{frame}

\subsection{Interface}

\begin{frame}{Using Landslide}
	\begin{center}
	\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	User must tell Landslide when certain kernel events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap
	% Make sure to expound on what these are
	Some complicated conditions require in-Landslide instrumentation.
	\begin{itemize}
		\item Scheduler design details
	\end{itemize}
\end{frame}

\begin{frame}{Decision Trace}
	% FIXME make the stack trace coloured
	\includegraphics[width=\textwidth]{found_a_bug.png}
	% Say: So landslide had to explore 5 different interleavings to find the bug. But this was a pretty simple test case, and 
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Addressing Challenges} % rename

\begin{frame}{State Space Reduction}
	% Say: ``combinatorial explosion''
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}
\begin{frame}{State Space Reduction}
	State spaces grow exponentially.
	\begin{itemize}
		\item Fortunately, some sequences result in identical states.
		\item Dynamic Partial Order Reduction\related{Flanagan '05}
		\begin{itemize}
			\item Requires ``memory independence relation'' between transitions.
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Shared Memory Conflicts} % slide: ignoring shared memory accesses - venn diagram
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict0.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict1.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict2.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	%\textbf{Problem:} Many conflicts we don't care about
	%\begin{itemize}
	%	\item Every transition accesses scheduler state.
	%	\item Many transitions access ``unrelated'' global objects.
	%\end{itemize}
	%\linegap
	\textbf{Solution:} Only consider ``relevant'' memory conflicts.
	\begin{itemize}
		\item Ignore scheduler, global objects assumed to be correct
		\begin{itemize}
			\item Sacrifice ability to test these to efficiently test everything else.
		\end{itemize}
		\item User must specify what to ignore.
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Identifying Bugs}
	\begin{columns}[l]
		\column{0.75\textwidth}
		\textbf{How do we know we've found a bug?}
		\linegap

		Definite bugs
		\begin{itemize}
			\item Kernel panics / assertion failures
			\item Use-after-free
			\item Deadlock
		\end{itemize}
		Probable bugs
		\begin{itemize}
			\item Infinite loop
			\begin{itemize}
				\item Use structure of execution tree to judge progress.
			\end{itemize}
		\end{itemize}
		\column{0.25\textwidth}
		\includegraphics[width=\textwidth]{blank-livelock.png}
	\end{columns}
\end{frame}
\begin{frame}[fragile]{Identifying Bugs}
	\begin{columns}[l]
		\column{0.75\textwidth}
		\textbf{How do we know we've found a bug?}
		\linegap

		Definite bugs
		\begin{itemize}
			\item Kernel panics / assertion failures
			\item Use-after-free
			\item Deadlock
		\end{itemize}
		Probable bugs
		\begin{itemize}
			\item Infinite loop
			\begin{itemize}
				\item Use structure of execution tree to judge progress.
			\end{itemize}
		\end{itemize}
		\column{0.25\textwidth}
		\includegraphics[width=\textwidth]{livelock.png}
	\end{columns}
\end{frame}
% slide: other_prob: decide on mutex_lock as opposed to everywhere, but...
% slide: manual shaping - have *user* specify what to focus on (automating intuition)
\begin{frame}{Focusing the Search Space}
	Recommended decision points: \texttt{mutex\_lock()} and \texttt{mutex\_unlock()}
	\linegap

	{\bf Problem:} Kernel uses mutexes {\em everywhere}.
	\begin{itemize}
		\item Good: \texttt{exit()} calls \texttt{mutex\_lock()}
		\item Bad: \texttt{exit()} calls \texttt{destroy\_address\_space()} calls \texttt{mutex\_lock()}
	\end{itemize}
	\linegap
	{\bf Solution:} User specifies what modules of the kernel to pay attention to.
	\begin{itemize}
		\item \texttt{within\_function exit}
		\item \texttt{without\_function destroy\_address\_space}
	\end{itemize}
\end{frame}
% slide: exciting teaser "i think this is not too far off!"
\begin{frame}{Teaser: (Semi-)Automated Search Focusing}
	Searching for bugs given a stack trace
	\begin{itemize}
		\item User knows {\em where} the bug is, but not {\em what}.
		\item Testing framework can use stack trace to conduct a more focused search.
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation}

\breakslide{Evaluating Landslide}

%\begin{frame}{Outline}
%	\done{Motivation: Concurrency debugging}
%	\begin{itemize}
%		\item \done{Systematic exploration versus stress testing}
%		\item \done{Challenges of kernel-space}
%	\end{itemize}
%	\linegap
%
%	\done{Tool: Landslide}
%	\begin{itemize}
%		\item \done{Design and interface}
%		\item \done{Addressing challenges}
%	\end{itemize}
%	\linegap
%
%	{\bf Evaluation: Finding Races}
%	\begin{itemize}
%		\item {\bf Student user study}
%		\item {\bf Case study}
%	\end{itemize}
%	\linegap
%
%	Future Work and Conclusion
%\end{frame}

\begin{frame}{Working with Students} % slide: worked with students - give some time breakdown
	Solicited students in 15-410 to use Landslide on their kernels.
	\begin{itemize}
		\item Can Landslide find bugs ``in the wild''?
		\item How much time does it take to get it working?
	\end{itemize}
	\linegap

	Five groups participated; four got it to work.
	\begin{itemize}
		\item Average instrumentation time 100 minutes (55 to 158)
		%\item Average customisation time 25 minutes (0 to 60)
	\end{itemize}
\end{frame}

\begin{frame}{Landslide Victories} % slide: anecdotes: two deterministic bugs & two races
	All four groups who got it to work found bugs.
	\linegap

	Race conditions
	\begin{itemize}
		\item Too many waiters allowed
		\begin{itemize}
			\item One parent thread might block forever
		\end{itemize}
		\item Accidental return to an exited thread
		% Also talk about their vanish/vanish bug
	\end{itemize}
	\linegap

	Deterministic bugs
	\begin{itemize}
		\item Inappropriate idling
		\begin{itemize}
			\item Waiting for a timer tick with no useful work to do
		\end{itemize}
		\item Use-after-free
		\begin{itemize}
			\item Required memory tracking to detect
		\end{itemize}
	\end{itemize}
\end{frame}

\subsection{Case Study}

\begin{frame}{Studying Bugs In-Depth} % slide: "I have built a software smarter than me!" 30-40sec to find
	Worked with two kernels: my own, and one I graded
	\linegap

	``I have made something smarter than me.''
	% Heh, if it were really smarter than me, I wouldn't have had to debug it.
	\begin{itemize}
		\item \texttt{double\_wait} - two parent threads \texttt{wait()} on one exiting child
		\begin{itemize}
			\item Expected: One thread reaps child, the other returns failure
			\item Got: \texttt{failed assertion `nobe->cond\_queue\_next == NULL'}
		\end{itemize}
		\item Previously unknown, not found by stress testing or by TA
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{frame}{Outline}
%	\done{Motivation: Concurrency debugging}
%	\begin{itemize}
%		\item \done{Systematic exploration versus stress testing}
%		\item \done{Challenges of kernel-space}
%	\end{itemize}
%	\linegap
%
%	\done{Tool: Landslide}
%	\begin{itemize}
%		\item \done{Design and interface}
%		\item \done{Addressing challenges}
%	\end{itemize}
%	\linegap
%
%	\done{Evaluation: Finding Races}
%	\begin{itemize}
%		\item \done{Student user study}
%		\item \done{Case study}
%	\end{itemize}
%	\linegap
%
%	{\bf Future Work and Conclusion}
%\end{frame}

%\subsection{Future Work}
\begin{frame}{Promising Directions}
	Top 15-410 students, effectively supervised, can find hard bugs faster.
	\begin{itemize}
		\item Use as a debugging tool may help students learn more.
	\end{itemize}
	\linegap

	After Pebbles\dots
	\begin{itemize}
		\item Potential application to production kernels
		\begin{itemize}
			\item General purpose OSes, device drivers
			% Be ready to field questions about SMP.
		\end{itemize}
	\end{itemize}
	\linegap

	Emphasis on users ``steering'' test by changing parameters
	\begin{itemize}
		\item Start with coarse granularity, iterate refining decision points
		\item Perhaps we can automate this steering?
	\end{itemize}
	\linegap
\end{frame}

%\subsection{Summary}
\begin{frame}{Summary}
	\textbf{Systematic testing in kernel-space}
	\begin{itemize}
		\item Use internal kernel abstractions to understand concurrency behaviour.
		\item Relying on user's knowledge makes testing easier.
	\end{itemize}
	\linegap
	\textbf{Landslide}
	\begin{itemize}
		\item Has potential to be a useful tool for 15-410.
		\item A first step towards sophisticated kernel debugging techniques.
	\end{itemize}

	\linegap
	\linegap
	\begin{center}
	{\scriptsize \em Please hold on. This transit is approaching the Landslide terminal, \\baggage claim, ground transportation, and ticketing.}
	\end{center}
\end{frame}

% slide: bonus slides

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

@@

@@

\begin{frame}{Execution Tree}
	\begin{center}
		%\includegraphics[width=0.9\textwidth]{threadfork0.png}
	\end{center}
\end{frame}
\begin{frame}{Execution Tree}
	\begin{center}
		%\includegraphics[width=0.9\textwidth]{threadfork05.png}
	\end{center}
\end{frame}
\begin{frame}{Execution Tree}
	\begin{center}
		%\includegraphics[width=0.9\textwidth]{threadfork1.png}
	\end{center}
\end{frame}
\begin{frame}{Execution Tree}
	\begin{center}
		%\includegraphics[width=0.9\textwidth]{threadfork2.png}
	\end{center}
\end{frame}

\begin{frame}{Decision Points} % : duplicate slide with later
	A {\bf decision point} is\ldots

@@

\begin{frame}{State Space Explosion}
	\begin{center}
	%\includegraphics[width=\textwidth]{undiamond0.png}
	\end{center}
\end{frame}
\begin{frame}{State Space Explosion}
	\begin{center}
	%\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}
\begin{frame}{State Space Explosion}
	\begin{center}
	%\includegraphics[width=\textwidth]{diamond1.png}
	\end{center}
\end{frame}

@@

\begin{frame}{Anatomy}
	\begin{center}
	%\includegraphics[width=0.6\textwidth]{landslide.png}
	\end{center}
\end{frame}

@@

\begin{frame}{Annotating Your Kernel}
	Step 1
	\begin{center}
		%\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	Your kernel needs to say when certain events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap

@@

\begin{frame}{Configuring Landslide}
	Step 2
	\begin{center}
		%\includegraphics[width=0.8\textwidth]{config-landslide.png}
	\end{center}
	Edit \texttt{config.landslide} with some details and tweaks

@@

@@

%\begin{frame}{Partial Order Reduction}
%	\begin{center}
%	The next 6 slides won't be on the exam.
%
%		\Huge \dbend
%	\end{center}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction}
%	A {\bf transition} is\ldots
%	\begin{itemize}
%		\item The sequence of instructions executed between choice points.
%	\end{itemize}
%	To express concurrency between transitions:
%	% Say: As a prerequisite for this algorithm, we need to build two sets to express the concurrency relations between the transitions
%	\begin{itemize}
%		\item Independence relation
%		\item Happens-before relation
%	\end{itemize}
%	% Say: This is hard, and I don't expect you to understand it during this lecture; you won't be tested on it or anything. My goal in presenting it is to convey how intensely cool the algorithm is, and hopefully you will keep thinking about it later. I'd be happy to answer questions about it later.
%\end{frame}
%
%\begin{frame}[fragile]{Partial Order Reduction - Independence Relation}
%	\begin{columns}[l]
%		\column{0.5\textwidth}
%		\begin{center}
%		\begin{verbatim}
%		void thread_A() {
%		    foo = 42;
%		}
%		void thread_B() {
%		    if (bar == 1337)
%		        print "Hello"
%		}
%		\end{verbatim}
%		\end{center}
%		\linegap
%
%		$A$ and $B$ are {\bf independent}.
%		\column{0.5\textwidth}
%		\begin{center}
%		\begin{verbatim}
%		void thread_A() {
%		    foo = 42;
%		}
%		void thread_C() {
%		    if (foo == 1337)
%		        print "Hello"
%		}
%		\end{verbatim}
%		\end{center}
%		\linegap
%
%		$A$ and $C$ {\bf conflict}.
%	\end{columns}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Independence Relation}
%	Consider a trace of all shared memory reads/writes
%
%	\linegap
%	Two transitions (instruction sequences) $A$ and $B$ are independent iff:
%	\begin{itemize}
%		\item $Writes(A) \cap Writes(B) = \emptyset$
%		\item $Writes(A) \cap Reads(B) = \emptyset$
%		\item $Reads(A) \cap Writes(B) = \emptyset$
%	\end{itemize}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Happens-before Relation}
%	\begin{columns}[l]
%		\column{0.5\textwidth}
%		%\includegraphics[width=\textwidth]{hb.png}
%		\column{0.5\textwidth}
%		Is a transition ``enabled'' by a previous one?
%		\begin{itemize}
%			\item $A$ enables $B$
%			\item $B$ enables $C$ (same thread)
%			\item $B$ enables $D$
%		\end{itemize}
%%		Transitive closure
%%		\begin{itemize}
%%			\item $HB^* = \{AB, AC, AD, BC, BD\}$
%%		\end{itemize}
%		$C$ and $D$ are {\bf concurrent}!
%	\end{columns}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Tagging}
%	Run POR at the end of each branch
%	\begin{itemize}
%		\item Identify which branches to explore next
%		\item Tag only a subset of all branches (hopefully)
%	\end{itemize}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Tagging}
%	For each transition $T$, find an {\bf evil ancestor} $E$:
%	\begin{itemize}
%		\item $E$ and $T$ are concurrent.
%		\item $E$ and $T$ are not independent.
%		\item Intuitively: $T$ and $E$ could be reordered, and might behave differently.
%	\end{itemize}
%	Find a {\bf good sibling} $G$ of $E$:
%	\begin{itemize}
%		\item $G$ is the same thread as $T$
%		% Say: (if no such sibling, all siblings)
%		\item Intuitively: Find a branch in which $T$ runs before $E$.
%	\end{itemize}
%	Explore tagged branches depth-first.
%\end{frame}
%
%\begin{frame}{Partial Order Reduction}
%	\begin{center}
%		{\Huge \dbend}
%		\linegap
%		\linegap
%		\linegap
%
%		Feel free to ask me about this later.
%	\end{center}
%\end{frame}

@@

% vim: foldmethod=indent

>>>>

\title[Landslide]{{\bf Systematic Dynamic Race Condition Detection}}
\subtitle[]{ {\em more clever than } ``\texttt{slaughter cho2}'' {\em since 2011.}}
\author[Ben Blum]{Ben Blum \texttt{(bblum@andrew.cmu.edu)}}

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2012, April 6}

@@

\usecolortheme[named=ForestGreen]{structure}

@@

\newcommand\linegap{\vspace{0.2in}}
\newcommand\breakslide[1]{\begin{frame}{} \begin{center} #1 \end{center} \end{frame}}

@@

\begin{frame}{Outline}
	\textbf{Theory: Seeing race conditions in a new way}
	\begin{itemize}
		\item Case study (example)
		\item The execution tree
		\item Decision points
	\end{itemize}
	{\bf Technique: Systematic testing}
	% Say: A way of enumerating all possible interleavings of a concurrent test case.
	\begin{itemize}
		\item Requirements
		\item Challenges and feasibility
	\end{itemize}
	{\bf Tool: Landslide}
	\begin{itemize}
		\item How it works
		\item The user-tool relationship
		\item User study (that's you!)
			% !!! SAY !!! : If you are struggling, this is not for you, etc.
	\end{itemize}
\end{frame}

@@

%	\begin{itemize}
%		\item On vanish, child's TCB is freed
%		\item Forking thread does use-after-free
%		\item Might return garbage instead of TID
%	\end{itemize}
\end{frame}

@@

\begin{frame}{Testing Mechanisms}
	\textbf{Stress testing}: \texttt{slaughter cho2} and friends
	\begin{itemize}
		\item Attempting to exercise as many interleavings as practical
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at just the right time\ldots''
		\end{itemize}
		\item Cryptic panic messages or machine reboots
	\end{itemize}
	What if\ldots
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Preempt enough times to run {\em every single} interleaving
		\item Tell the story of what {\em actually happened}.
		\item Overlook fewer bugs!
	\end{itemize}
	% Say: At the end of this lecture, I'm going to be looking for some among you...
\end{frame}

@@

\begin{frame}{Execution Tree}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{threadfork0.png}
	\end{center}
\end{frame}
\begin{frame}{Execution Tree}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{threadfork05.png}
	\end{center}
\end{frame}
\begin{frame}{Execution Tree}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{threadfork1.png}
	\end{center}
\end{frame}
\begin{frame}{Execution Tree}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{threadfork2.png}
	\end{center}
\end{frame}

\begin{frame}{Decision Points} % TODO: duplicate slide with later
	A {\bf decision point} is\ldots

@@

\begin{frame}{State Space Explosion}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond0.png}
	\end{center}
\end{frame}
\begin{frame}{State Space Explosion}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}
\begin{frame}{State Space Explosion}
	\begin{center}
	\includegraphics[width=\textwidth]{diamond1.png}
	\end{center}
\end{frame}

@@

\begin{frame}{Anatomy}
	\begin{center}
	\includegraphics[width=0.6\textwidth]{landslide.png}
	\end{center}
\end{frame}

@@

\begin{frame}{Annotating Your Kernel}
	Step 1
	\begin{center}
		\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	Your kernel needs to say when certain events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap

@@

\begin{frame}{Configuring Landslide}
	Step 2
	\begin{center}
		\includegraphics[width=0.8\textwidth]{config-landslide.png}
	\end{center}
	Edit \texttt{config.landslide} with some details and tweaks

@@

% TODO more of these with screenshots

@@

\end{document}
%\begin{frame}{Partial Order Reduction}
%	\begin{center}
%	The next 6 slides won't be on the exam.
%
%		\Huge \dbend
%	\end{center}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction}
%	A {\bf transition} is\ldots
%	\begin{itemize}
%		\item The sequence of instructions executed between choice points.
%	\end{itemize}
%	To express concurrency between transitions:
%	% Say: As a prerequisite for this algorithm, we need to build two sets to express the concurrency relations between the transitions
%	\begin{itemize}
%		\item Independence relation
%		\item Happens-before relation
%	\end{itemize}
%	% Say: This is hard, and I don't expect you to understand it during this lecture; you won't be tested on it or anything. My goal in presenting it is to convey how intensely cool the algorithm is, and hopefully you will keep thinking about it later. I'd be happy to answer questions about it later.
%\end{frame}
%
%\begin{frame}[fragile]{Partial Order Reduction - Independence Relation}
%	\begin{columns}[l]
%		\column{0.5\textwidth}
%		\begin{center}
%		\begin{verbatim}
%		void thread_A() {
%		    foo = 42;
%		}
%		void thread_B() {
%		    if (bar == 1337)
%		        print "Hello"
%		}
%		\end{verbatim}
%		\end{center}
%		\linegap
%
%		$A$ and $B$ are {\bf independent}.
%		\column{0.5\textwidth}
%		\begin{center}
%		\begin{verbatim}
%		void thread_A() {
%		    foo = 42;
%		}
%		void thread_C() {
%		    if (foo == 1337)
%		        print "Hello"
%		}
%		\end{verbatim}
%		\end{center}
%		\linegap
%
%		$A$ and $C$ {\bf conflict}.
%	\end{columns}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Independence Relation}
%	Consider a trace of all shared memory reads/writes
%
%	\linegap
%	Two transitions (instruction sequences) $A$ and $B$ are independent iff:
%	\begin{itemize}
%		\item $Writes(A) \cap Writes(B) = \emptyset$
%		\item $Writes(A) \cap Reads(B) = \emptyset$
%		\item $Reads(A) \cap Writes(B) = \emptyset$
%	\end{itemize}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Happens-before Relation}
%	\begin{columns}[l]
%		\column{0.5\textwidth}
%		\includegraphics[width=\textwidth]{hb.png}
%		\column{0.5\textwidth}
%		Is a transition ``enabled'' by a previous one?
%		\begin{itemize}
%			\item $A$ enables $B$
%			\item $B$ enables $C$ (same thread)
%			\item $B$ enables $D$
%		\end{itemize}
%%		Transitive closure
%%		\begin{itemize}
%%			\item $HB^* = \{AB, AC, AD, BC, BD\}$
%%		\end{itemize}
%		$C$ and $D$ are {\bf concurrent}!
%	\end{columns}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Tagging}
%	Run POR at the end of each branch
%	\begin{itemize}
%		\item Identify which branches to explore next
%		\item Tag only a subset of all branches (hopefully)
%	\end{itemize}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Tagging}
%	For each transition $T$, find an {\bf evil ancestor} $E$:
%	\begin{itemize}
%		\item $E$ and $T$ are concurrent.
%		\item $E$ and $T$ are not independent.
%		\item Intuitively: $T$ and $E$ could be reordered, and might behave differently.
%	\end{itemize}
%	Find a {\bf good sibling} $G$ of $E$:
%	\begin{itemize}
%		\item $G$ is the same thread as $T$
%		% Say: (if no such sibling, all siblings)
%		\item Intuitively: Find a branch in which $T$ runs before $E$.
%	\end{itemize}
%	Explore tagged branches depth-first.
%\end{frame}
%
%\begin{frame}{Partial Order Reduction}
%	\begin{center}
%		{\Huge \dbend}
%		\linegap
%		\linegap
%		\linegap
%
%		Feel free to ask me about this later.
%	\end{center}
%\end{frame}

% TODO: run through an example

@@
